A simple and efficient algorithm for computing approximate Nash equilibria in two-person zero-sum sequential games with imcomplete information

### 2015-07-23 paper.tex
fixups

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e}
%\documentclass[envcountsame]{llcns2e/llncs}
\usepackage{url}
\usepackage{times}
\usepackage{amsmath,amsfonts, amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage[FIGTOPCAP]{subfigure}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplexes]{
    \includegraphics[width=.5\linewidth]{0.pdf}
  }
  %% \hspace{-2em}
  %% \subfigure[Toy Poker]{
  %%   \includegraphics[width=.5\linewidth]{SimplifiedPoker_dgap.pdf}
  %% }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.5\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-2em}
  \includegraphics[width=.5\linewidth]{1.pdf}
  %% \hspace{-1em}
  %% \includegraphics[width=.5\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.5\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: Evolution of ergodic
    primal-dual gap. \textbf{Bottom row}: Evolution of
    value of game.}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{Experiments on sequential games with imcomplete information: Kuhn 3-card
  poker.} This game is a simplified form of poker
developed by Harold W. Kuhn. It is a two-person zero-sum game which is
simply enough to serve as an example but contains all the complexity
traits of a imcomplete information sequential game. The
deck includes only three playing cards, for example a King, Queen, and
Jack. One card is dealt to each player, then the first player must bet
or pass, then the second player may bet or pass. If any player chooses
to bet the opposing player must bet as well ("call") in order to stay
in the round. After both players pass or bet the player with the
highest card wins the pot. The
sequence-form representation of the game has $n_1 = n_2 = 13$
sequences and $l_1 = l_2 = 7$ information
sets per player. The pair $(x^*, y^*) \in \mathbb{R}^{13} \times
\mathbb{R}^{13}$ of realization plans given by
\begin{eqnarray*}
  \begin{split}
    x^* &= [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0,
      0.275, 0.725]^T,\\
    y^* &= [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T
    \end{split}
\end{eqnarray*}
is a Nash $10^{-4}$-equlibrium computed in 1500 iterations of
Algorithm  \ref{Tab:algo}. The convergence curves are shown
in Fig \ref{Tab:dgap_curve}. One easy checks that this equilibrium is
feasible. Indeed, $E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67
      \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5},
      -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$
and
$E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
  \times 10^{-6}, -1.50 \times 10^{-6},
      2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T$.
Finally, one checks that ${x^*}^TAy^* = {-0.05555}$,
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. The
 evolution of the dual gap and the expected value of the game across
 iterations are shown in Figure \ref{Tab:dgap_curve}.
More details (payoff matrix, etc.) are provided in
the supplementary materials...

\section{Concluding remarks}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a simple and efficient primal-dual algorithm for computing
Nash equilibria in two-person zero-sum sequential games with
imcomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a very low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game. Equilibrium problems are
saddle-point convex-concave problems, and as
such a natural choice for algorithms for solving them would be in the
family of primal-dual algorithms. We believe such primal-dual
schemes will receive more attention in the algorithmic game theory
community in future.

>>>>

\documentclass{article} % For LaTeX2e\usepackage{nips14submit_e}
%\documentclass[envcountsame]{llcns2e/llncs}
\usepackage{url}
\usepackage{times}
\usepackage{amsmath,amsfonts, amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmicx}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}
\usepackage{algpseudocode}
\usepackage{pifont}
\usepackage[FIGTOPCAP]{subfigure}

@@

\begin{figure}[!htpb]
  \subfigure[Matrix game on simplexes]{
    \includegraphics[width=.5\linewidth]{0.pdf}
  }
  %% \hspace{-2em}
  %% \subfigure[Toy Poker]{
  %%   \includegraphics[width=.5\linewidth]{SimplifiedPoker_dgap.pdf}
  %% }
  \hspace{-2em}
  \subfigure[Kuhn 3-card Poker]{
    \includegraphics[width=.5\linewidth]{Kuhn3112_dgap.pdf}
  }
  \vspace{-1em}
  \hspace{-2em}
  \includegraphics[width=.5\linewidth]{1.pdf}
  %% \hspace{-1em}
  %% \includegraphics[width=.5\linewidth]{SimplifiedPoker_NE.pdf}
  \hspace{-1em}
  \includegraphics[width=.5\linewidth]{Kuhn3112_NE.pdf}
  \caption{Convergence curves of Algorithm
    \ref{Tab:algo}. \textbf{Top row}: Evolution of ergodic
    primal-dual gap. \textbf{Bottom row}: Evolution of
    value of game...}
  \label{Tab:dgap_curve}
\end{figure}

\paragraph{Experiments on sequential games with imcomplete information: Kuhn 3-card
  poker.} This game is a simplified form of poker
developed by Harold W. Kuhn. It is a two-person zero-sum game which is
simply enough to serve as an example but contains all the complexity
traits of a imcomplete information sequential game. The
deck includes only three playing cards, for example a King, Queen, and
Jack. One card is dealt to each player, then the first player must bet
or pass, then the second player may bet or pass. If any player chooses
to bet the opposing player must bet as well ("call") in order to stay
in the round. After both players pass or bet the player with the
highest card wins the pot. The
sequence-form representation of the game has $n_1 = n_2 = 13$
sequences and $l_1 = l_2 = 7$ information
sets per player. The pair $(x^*, y^*) \in \mathbb{R}^{13} \times
\mathbb{R}^{13}$ of realization plans given by
\begin{eqnarray*}
  \begin{split}
    x^* &= [1, 0.759, 0.759, 0, 0.241, 1, 0.425, 0.575, 0, 0.275, 0,
      0.275, 0.725]^T,\\
    y^* &= [1, 1, 0, 0.667, 0.333, 0.667, 0.333, 1, 0, 0, 1, 0, 1]^T
    \end{split}
\end{eqnarray*}
is a Nash $10^{-4}$-equlibrium computed in 1500 iterations of
Algorithm  \ref{Tab:algo}. The convergence curves are shown
in Fig \ref{Tab:dgap_curve}. One easy checks that this equilibrium is
feasible. Indeed, $E_1x^* - e_1 = [4.76 \times 10^{-5}, -1.91 \times 10^{-5}, 5.67
      \times 10^{-5}, 8.23 \times 10^{-6}, 2.90 \times 10^{-5},
      -8.62 \times 10^{-7}, -1.96 \times 10^{-5}]^T$
and
$E_2y^* - e_2 = [-7.04 \times 10^{-7}, 2.27 \times 10^{-6}, -3.29
  \times 10^{-6}, -1.50 \times 10^{-6},
      2.92 \times 10^{-6}, -4.97 \times 10^{-7}, -5.85 \times
      10^{-7}]^T$.
Finally, one checks that ${x^*}^TAy^* = {-0.05555}$,
 which agrees to 5 d.p with the value of $-1 / 18$ computed
 analytically by H. W. Kuhn in his 1950 paper \cite{kuhn}. More
 details (payoff matrix, etc.) are provided in the supplementary
 materials. The  evolution of the dual gap and the expected value of
 the game across iterations are shown in Figure \ref{Tab:dgap_curve}.
We have not benchmarked this against the algorithms proposed in
\cite{nesterov2005a} and Gilpin's et al. \cite{gilpinfirst} because
implementing them for such games would require us to compute the
complicated projections $\prox_{Q_k}$.  We recall that
avoiding these projections was one of the goals of the present
manuscript.

\section{Concluding remarks}
Making use of the sequence-form representation
\cite{koller1992complexity,von1996efficient,vonequilibrium}, we have
deviced a simple and efficient primal-dual algorithm for computing
Nash equilibria in two-person zero-sum sequential games with
imcomplete information (like Texas Hold'em, etc.). Our algorithm is
simple to implement, with a low constant cost per iteration, and
enjoys a rigorous convergence theory with a proven
$\mathcal{O}(1/\epsilon)$ convergence in terms of basic operations
(matvec products, clipping, etc.), to a Nash
$(\epsilon,0)$-equilibrium of the game. Equilibrium problems are
saddle-point convex-concave problems, and as
such a natural choice for algorithms for solving them would be in the
family of primal-dual algorithms. We believe such primal-dual
schemes will receive more attention in the algorithmic game theory
community in future.
